---
title: 'Data Analysis 3: Week 10'
author: "Alexey Bessudnov"
date: "21 March 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Plan for today.

- Assignment 5: discuss
- Functions in R

Exercises.

Exercise 1. Write a function to calculate the mean of a numeric vector.

```{r}
myMean <- function(x) {
  sum(x) / length(x)
}

myMean(1:10)
mean(1:10)

```

Exercise 2. Modify this function to include an extra argument to deal with missing values.

```{r}
myMean2 <- function(x, rm.missing = FALSE) {
        if (rm.missing == TRUE) {
                x <- na.omit(x)
        }
        sum(x) / length(x)
}

myMean(1:10)
myMean2(c(1:10, NA))
myMean2(c(1:10, NA), rm.missing = TRUE)
```

Exercise 3. Modify this function to return an error when the vector is not numeric.

```{r}
myMean3 <- function(x, rm.missing = FALSE) {
        if (!is.numeric(x)) {
                stop("The vector is not numeric.")
        }
        if (rm.missing == TRUE) {
                x <- na.omit(x)
        }
        sum(x) / length(x)
}
# myMean3(c(1:10, "a"))

```

Exercise 4. Modify this function so that it saves the mean in the environment with the name "meanx". (Hint: think about environments.)

```{r}
myMean(1:10)
# meanx <- myMean(1:10)
myMean4 <- function(x) {
        meanx <<- sum(x) / length(x)
        meanx
}
myMean4(1:10)
```

Exercise 5. Write a function to calculate the factorial (i.e. 5! = 1x2x3x4x5). Note that 0! = 1, and for the negative numbers the factorial is not defined. (Hint: use recursion.)

```{r}
myFactorial <- function(x){
  if (x == 0) {return(1)}
  x * myFactorial(x-1)
}

myFactorial(4)

```

Exercise 6. Use this function to calculate the factorial for the numbers from 1 to 10. (Hint: use map()).

```{r}
# this is not going to work because the function has not been vectorised
myFactorial(1:10)

# a for() loop
output <- numeric(10)
for (i in 1:10) {
  output[i] <- myFactorial(i)
}
output

# with map()

library(tidyverse)

mapply(myFactorial, 1:10)
map_dbl(1:10, myFactorial)

```


Exercise 7. Open the indresp file for wave 8 and keep two variables: sex (h_sex_dv) and total net personal income (h_fimnnet_dv). You may need to clean sex and retain only values for male and female. Keep only cases with non-missing values for sex and income.

a) regress income on sex (i.e. find the difference between mean income for men and women). Store the estimate for the regression coefficient.
b) Now regress income on sex 100 times, each time with a random sample of 1,000 people. Store 100 regression coefficients in a vector and illustrate their distribution graphically.
c) Now do the same, changing the sample size to 100 people, and compare two distributions.

```{r}
df <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab") %>%
        select(h_sex_dv, h_fimnnet_dv)

df <- df %>%
        mutate(h_sex_dv = ifelse(h_sex_dv == 1, "male",
                      ifelse(h_sex_dv == 2, "female", NA))) %>%
        filter(!is.na(h_sex_dv)) %>%
        filter(!is.na(h_fimnnet_dv))

est <- lm(h_fimnnet_dv ~ h_sex_dv, df)$coefficients[[2]]

output <- numeric(100)
for (i in 1:100) {
  df1 <- df %>% sample_n(1000)
  output[i] <- lm(h_fimnnet_dv ~ h_sex_dv, df1)$coefficients[[2]]
}
output

sexCoef <- map(1:100, ~ df %>% sample_n(100)) %>%
        map_dbl(~ lm(h_fimnnet_dv ~ h_sex_dv, data = .)$coefficients[2])

enframe(sexCoef) %>%
        ggplot(aes(x = value)) +
        geom_histogram() +
        geom_vline(xintercept = est, colour = "red")

```


