---
title: 'Data Analysis 3: Class 3'
author: "Alexey Bessudnov"
date: "31 January 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

Plan for today:

1. Test assignment.
2. The **tidyverse** framework (https://www.tidyverse.org/).
3. Reading in data with **readr**.
4. Transforming data with **dplyr**.
5. Statistical assignment 1: questions.
6. Homework for next week.

Importing data: read ch.11 from R for Data Science (Data import): https://r4ds.had.co.nz/data-import.html and ch.2 from my website (Read data): http://abessudnov.net/dataanalysis3/readdata.html.

```{r}
library(tidyverse)
Data <- read_tsv("data/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
```

This is a cross-wave data file with stable characteristics of individuals. See the codebook at https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/wave/xwave/datafile/xwavedat.

See the dplyr cheetsheet: https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf

Exercises.

1. Select the variables for: sex (derived), date of birth (derived), ethnic group (racel_dv). Also keep the cross-wave identifier (pidp) and the sample origin variable (memorig).

```{r}

# subset(Data, select = c("pidp", "memorig", "sex_dv", "doby_dv", "racel_dv"))
# Data[,c(1, 5, 15, 19)]

select(Data, pidp, memorig, sex_dv, doby_dv, racel_dv)

# table(Data$sex_dv)
# prop.table(table(Data$sex_dv))
# prop.table(table(Data$sex_dv)) * 100
# 
# Data$sex_dv %>%
#   table() %>%
#   prop.table()

Data <- Data %>%
  select(pidp, memorig, sex_dv, doby_dv, racel_dv)

```

2. Filter the data to keep (in new data frames):
        a) men only.
        b) people born before 1950 and after 1975.
        c) men of Pakistani origin born in 1958 or 1982. 

Discuss: logical operators.

```{r}
Data %>%
  filter(sex_dv == 1)

Data %>%
  filter(doby_dv > 0 & (doby_dv < 1950 |  doby_dv > 1975))

Data %>%
  filter(sex_dv == 1 & racel_dv == 10 & (doby_dv == 1958 | doby_dv == 1982))

# Extra:

Data %>%
  filter(doby_dv %in% 1950:1974)

```


3. Recode birth year into cohorts (a new variable): the G.I. Generation (born before 1924), Silent Generation (1925-42), Baby Boomers (1943-65), Generation X (1966-1980), Millenials (1981-99), Generation Z (2000-). (The years are approximate.)

```{r}
Data <- Data %>%
  mutate(generation = case_when(
    between(doby_dv, 0, 1924) ~ "GI Generation",
    between(doby_dv, 1925, 1942) ~ "Silent Generation",
    between(doby_dv, 1943, 1965) ~ "Baby Boomers",
    between(doby_dv, 1966, 1980) ~ "Generation X",
    between(doby_dv, 1981, 1999) ~ "Millenials",
    doby_dv >= 2000 ~ "Generation Z"
  ))

Data %>%
  pull(generation) %>%
  class()

Data %>%
  pull(generation) %>%
  table()


```

4. Recode ethnicity into the following groups: white British, Other White, Indian, Pakistani, other. (This classification doesn't make much sense, but we're doing this as an exercise).

Discuss: numeric and character vectors (strings), factors. Missing values in R.

5. Count the number of people belonging to different ethnic groups (and produce percentages). 

```{r}
Data %>%
  count(racel_dv) %>%
  mutate(perc = n / sum(n) * 100)

```

6. Summarise the proportion of white British by generation.

```{r}
Data %>%
  filter(racel_dv != -9) %>%
  mutate(whiteBritish = if_else(racel_dv == 1, 1, 0)) %>%
  group_by(generation) %>%
  summarise(
    propWhiteBritish = mean(whiteBritish, na.rm = TRUE) * 100
  )

```


7. Summarise the proportion of women by birth year.














