---
title: 'Data Analysis 3: Class 2'
author: "Alexey Bessudnov"
date: "24 January 2019"
output:
  ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

## Plan for today

- Assignments and submission
- Reproducible research, project workflow, R Markdown
- Understanding Society Data structure
- Reading data in R

## Assignments

- Test assignment (deadline 29 Jan, 2pm): https://github.com/dataanalysis3/testAssignment2019
- Accept the invitation that I sent by email and complete the assignment using Github Classroom
- Once completed do not forget to submit on Bart
- Assignment 1 (deadline 5 February, 2pm): https://github.com/dataanalysis3/assignment1

## Reproducible research

- It should be possible to reproduce every statistical analysis you've done  (by other researchers or future you)
- Always keep your code in scripts
- Have a folder for each project (R Studio project) and create subfolders with a clear structure
- Always write comments for your code so that other people (or future you) can understand what you have done
- Organise your code in a clear way and follow style conventions (see, for example, Google's R Style Guide: https://google.github.io/styleguide/Rguide.xml)
- Use Github

## R Markdown

- Combine code and output in one document
- Easily change output formats (html, LaTeX, Word, etc.)
- See [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)
- Also see http://abessudnov.net/dataanalysis3/rmarkdown.html
- Alternative: Jupyter Notebook
- Now let us explore how it works

## Exercise 1

- Complete problems 1, 2 and 3 from your test statistical assignment
- Experiment with different output formats
- Commit and push the .Rmd and .md files on Github (do not commit files in other output formats)

## Understanding Society data structure

- Read the User Guide, pp. 57-64: https://www.understandingsociety.ac.uk/sites/default/files/downloads/documentation/mainstage/user-guides/mainstage-user-guide.pdf
- Let us explore data files
- **mrdoc** contains all the documentation 
- **tab** contains data files in the tabulated format

## Data structure 2

- 18 waves of the BHPS data
- 8 waves of UndSoc data + cross-wave data
- Codebooks available here: https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation
- Let us look at the latest wave 8 and explore the data files

## Reading data in R

- You can read data in R in several ways
- Base R (*read.csv()* etc.)
- We will use the **readr** package (part of **tidyverse**)
- For tab separated files we want the function *read_tsv()*
- At a later point, we will use the *fread()* function from the **data.table** package


## Wave 8: data from the individual adult questionnaire (h_indresp)

```{r}
# attaching the tidyverse library quietly
suppressMessages(library(tidyverse))

indresp <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")

# remove the data to free up memory
rm(indresp)

```

## Household-level substantive file (h_hhresp)


```{r}
hhresp <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_hhresp.tab")

rm(hhresp)

```

## Household roster data (h_egoalt)

```{r}
egoalt <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_egoalt.tab")

rm(egoalt)

```

## Stable individual characteristics

```{r}
xwavedat <- read_tsv("data/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")

rm(xwavedat)

```

## Other data files

- h_child: data on children
- h_youth: data from self-completion questionnaire for children
- h_indall: all people in the household, inc. children and non-responses
- h_indsamp: technical individual-level data
- h_hhsamp: technical and observational data on hoseholds
- h_callrec: call records
- h_income: income and payments individual level data
- h_newborn: data on newborn children
- h_parstyle: parenting style data
