---
title: 'Data Analysis 3: Week 6'
author: "Alexey Bessudnov"
date: "21 February 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)

```
Plan for today:

- Assignment 2: solution
- Assignment 3
- Simple data visualisation 
- Homework for next week

## Exercises

1. Read the *indresp* file from Wave 8 and keep the following variables: pidp, derived sex and age, ethnic group (h_indresp), government office region (h_gor_dv), and net personal income (h_fimnnet_dv).

    ```{r}
library(tidyverse)
Data8 <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
Data8 <- Data8 %>%
  select(pidp, h_sex_dv, h_age_dv, h_gor_dv, h_fimnnet_dv)

    ```

For all charts use ggplot2. You may need to clean and recode variable before visualising.

We will start with univariate distributions.

2. Visualise the distribution of income with a histogram, a density plot and a box plot.

    ```{r}
ggplot(Data8,
       aes(x = h_fimnnet_dv)) +
  geom_histogram(binwidth = 100) +
  xlim(-100, 5000) + 
  xlab("Net monthly personal income")

ggplot(Data8,
       aes(x = h_fimnnet_dv)) +
  geom_density() +
  xlim(-100, 5000)

ggplot(Data8,
       aes(y = h_fimnnet_dv)) +
  geom_boxplot() +
  ylim(-100,5000)

    ```

3. Visualise the distribution of sex with a bar graph. 

    ```{r}
Data8 %>%
  mutate(h_sex_dv = ifelse(h_sex_dv == 1, "male",
                           ifelse(h_sex_dv == 2, "female", NA))) %>%
  filter(!is.na(h_sex_dv)) %>%
  ggplot(
    aes(x = h_sex_dv)
  ) +
  geom_bar() +
  xlab("Sex")

# ggplot(Data8,
#        aes(x = h_sex_dv)) +
#   geom_bar()


Data8 %>%
  mutate(h_sex_dv = ifelse(h_sex_dv == 1, "male",
                           ifelse(h_sex_dv == 2, "female", NA))) %>%
  filter(!is.na(h_sex_dv)) %>%
  ggplot(
    aes(x = h_sex_dv)
  ) +
  geom_bar(aes(y = (..count..)/sum(..count..) * 100)) +
  xlab("Sex") +
  ylab("Percent") +
  scale_y_continuous(breaks=seq(0, 60, 10))


    ```

Bivariate distributions.

4. Create a bar chart showing mean income by region.

    ```{r}
Data8 <- Data8 %>%
  mutate(region = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "Souh West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland"))
byRegion <- Data8 %>%
  filter(!is.na(region)) %>%
  group_by(region) %>%
  summarise(
    medianIncome = median(h_fimnnet_dv, na.rm = TRUE)
  )

byRegion %>%
ggplot(
  aes(x = reorder(region, medianIncome), y = medianIncome)
  ) +
  geom_bar(stat = "identity") +
  xlab("") +
  ylab("Median net monthly personal income") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

    ```


5. Make a dot plot showing the same information as above (without splitting by sex). Sort regions in the descending order by mean income.

    ```{r}
byRegion %>%
ggplot(
  aes(y = reorder(region, medianIncome), x = medianIncome)
  ) +
  geom_point(size = 3) +
  xlab("Median net monthly personal income") +
  ylab("")
    ```

6. Make a line chart showing median income by age.

    ```{r}
byAge <- Data8 %>%
  group_by(h_age_dv) %>%
  summarise(
    medianIncome = median(h_fimnnet_dv, na.rm = TRUE)
  )

byAge %>%
        ggplot(aes(x = h_age_dv, y = medianIncome)) +
        geom_line() +
        geom_smooth() +
        xlim(21,80) +
        xlab("Age") +
        ylab("Median income")

# Split by sex

byAgeSex <- Data8 %>%
  mutate(h_sex_dv = ifelse(h_sex_dv == 1, "male",
                           ifelse(h_sex_dv == 2, "female", NA))) %>%
  filter(!is.na(h_sex_dv)) %>%
  group_by(h_age_dv, h_sex_dv) %>%
  summarise(
    medianIncome = median(h_fimnnet_dv, na.rm = TRUE)
  )

byAgeSex %>%
        ggplot(aes(x = h_age_dv, y = medianIncome, colour = h_sex_dv)) +
        geom_line() +
        geom_smooth() +
        xlim(21,80) +
        ylim(0,2500) +
        xlab("Age") +
        ylab("Median income")
    ```

