---
title: 'Data Analysis 3: Class 3'
author: "Alexey Bessudnov"
date: "31 January 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

Plan for today:

1. Test assignment.
2. The **tidyverse** framework (https://www.tidyverse.org/).
3. Reading in data with **readr**.
4. Transforming data with **dplyr**.
5. Statistical assignment 1: questions.
6. Homework for next week.

Importing data: read ch.11 from R for Data Science (Data import): https://r4ds.had.co.nz/data-import.html and ch.2 from my website (Read data): http://abessudnov.net/dataanalysis3/readdata.html.

```{r}
library(tidyverse)
Data <- read_tsv("data/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
```

This is a cross-wave data file with stable characteristics of individuals. See the codebook at https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/wave/xwave/datafile/xwavedat.

See the dplyr cheetsheet: https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf

Exercises.

1. Select the variables for: sex (derived), date of bith (derived), ethnic group (racel_dv). Also keep the cross-wave identifier (pidp) and the sample origin variable (memorig). 

2. Filter the data to keep (in new data frames):
        a) men only.
        b) people born before 1950 and after 1975.
        c) men of Pakistani origin born in 1958 or 1982. 

Discuss: logical operators.

3. Recode birth year into cohorts (a new variable): the G.I. Generation (born before 1924), Silent Generation (1925-42), Baby Boomers (1943-65), Generation X (1966-1980), Millenials (1981-99), Generation Z (2000-). (The years are approximate.)

4. Recode ethnicity into the following groups: white British, Other White, Indian, Pakistani, other. (This classification doesn't make much sense, but we're doing this as an exercise).

Discuss: numeric and character vectors (strings), factors. Missing values in R.

5. Count the number of people belonging to different ethnic groups (and produce percentages). 

6. Summarise the proportion of white British by generation.

7. Summarise the proportion of women by birth year.














