---
title: 'Data Analysis 3: Class 3'
author: "Alexey Bessudnov"
date: "31 January 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

Plan for today:

- Test assignment.
- The **tidyverse** framework (https://www.tidyverse.org/).
- Reading in data with **readr**.
- Transforming data with **dplyr**.
- Statistical assignment 1: questions.
- Homework for next week.

Importing data: read ch.11 from R for Data Science (Data import): https://r4ds.had.co.nz/data-import.html and ch.2 from my website (Read data): http://abessudnov.net/dataanalysis3/readdata.html.

```{r}
library(tidyverse)
Data <- read_tsv("data/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
```

This is a cross-wave data file with stable characteristics of individuals. See the codebook at https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/wave/xwave/datafile/xwavedat.

See the dplyr cheetsheet: https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf

Exercises.

1. Select the variables for: sex (derived), date of birth (derived), ethnic group (racel_dv). Also keep the cross-wave identifier (pidp) and the sample origin variable (memorig).

    ```{r}

# With base R:
subset(Data, select = c("pidp", "memorig", "sex_dv", "doby_dv", "racel_dv"))
Data[,c(1, 5, 15, 19)]

# with dplyr
select(Data, pidp, memorig, sex_dv, doby_dv, racel_dv)

# using a pipe (%>%)

Data <- Data %>%
  select(pidp, memorig, sex_dv, doby_dv, racel_dv)

    ```

2. Filter the data to keep (in new data frames):
        a) men only.
        b) people born before 1950 and after 1975.
        c) men of Pakistani origin born in 1958 or 1982. 


    ```{r}
# a)
Data %>%
  filter(sex_dv == 1)

# b)
Data %>%
  filter(doby_dv > 0 & (doby_dv < 1950 |  doby_dv > 1975))

# c)
Data %>%
  filter(sex_dv == 1 & racel_dv == 10 & (doby_dv == 1958 | doby_dv == 1982))


    ```

3. Recode birth year into cohorts (a new variable): the G.I. Generation (born before 1924), Silent Generation (1925-42), Baby Boomers (1943-65), Generation X (1966-1980), Millenials (1981-99), Generation Z (2000-). (The years are approximate.)

    ```{r}
# with case_when() (good with complex conditions)
Data <- Data %>%
  mutate(generation = case_when(
    between(doby_dv, 0, 1924) ~ "GI Generation",
    between(doby_dv, 1925, 1942) ~ "Silent Generation",
    between(doby_dv, 1943, 1965) ~ "Baby Boomers",
    between(doby_dv, 1966, 1980) ~ "Generation X",
    between(doby_dv, 1981, 1999) ~ "Millenials",
    doby_dv >= 2000 ~ "Generation Z"
  ))
# case_when is particularly useful when there are multiple variables in logical statements, for example:

# men, 18 to 25 years old, recoded to "young men"
# (the variable names do not refer to our data; this is just an example.)
# case_when(
#         between(age, 18, 25) & sex == "male" ~ "young men"
# )

# You can also use ifelse for recoding, but it works best with simple cases.
# recode sex into "male" or "female"", dropping other cases.

Data %>%
        mutate(sexBinary = ifelse(sex_dv == 1, "male",
                                   ifelse(sex_dv == 2, "female", NA)))

    ```

4. Recode ethnicity into the following groups: white British, Other White, Indian, Pakistani, other. (This classification doesn't make much sense, but we're doing this as an exercise).


    ```{r}
table(Data$racel_dv)

# using recode

Data <- Data %>%
        mutate(ethnRecoded = recode(racel_dv,
                `1` = "white British",
                `2` = "other White",
                `3` = "other White",
                `4` = "other White",
                `9` = "Indian",
                `10` = "Pakistani",
                `-9` = NA_character_,
                # .default is for all other values
                .default = "other"
        )) %>%
        # let's make this new variable a factor
        mutate(ethnRecoded = factor(ethnRecoded))

# checking that the recoding was correct
Data %>%
        count(racel_dv, ethnRecoded)

# we could do the same recoding with case_when

Data %>%
        mutate(ethnRecoded = case_when(
                racel_dv == 1 ~ "white British",
                between(racel_dv, 2, 4) ~ "other White",
                racel_dv == 9 ~ "Indian",
                racel_dv == 10 ~ "Pakistani",
                (racel_dv >= 5 & racel_dv <= 8) | racel_dv > 10 ~ "other",
                racel_dv == -9 ~ NA_character_
        )) %>%
        count(racel_dv, ethnRecoded)
        

    ```


5. Count the number of people belonging to different ethnic groups (and produce percentages). 

    ```{r}
Data %>%
  count(ethnRecoded) %>%
  mutate(perc = n / sum(n) * 100)

    ```

6. Summarise the proportion of white British by generation.

    ```{r}
Data %>%
  filter(racel_dv != -9) %>%
  mutate(whiteBritish = if_else(racel_dv == 1, 1, 0)) %>%
  group_by(generation) %>%
  summarise(
    propWhiteBritish = mean(whiteBritish, na.rm = TRUE) * 100
  )

    ```

7. Summarise the percentage of women by birth year.

    ```{r}
Data %>%
        filter(doby_dv > 0) %>%
        mutate(female = ifelse(sex_dv == 2, TRUE,
                                   ifelse(sex_dv == 1, FALSE, NA))) %>%
        group_by(doby_dv) %>%
        summarise(
                propFemale = mean(female, na.rm = TRUE) * 100
                )

    ```












